---
title: 'MonthlyStudyPermitHolders: Cntry & Qrts'
author: "McKenzie"
date: "2023-12-09"
output: html_document
---

### Packages
```{r setup, include=FALSE}

library(readxl)
library(dplyr)
library(tidyr)
library(openxlsx)
library(httr)
```

### Read Data Source

```{r}
URL<-'https://www.ircc.canada.ca/opendata-donneesouvertes/data/EN_ODP-TR-Study-IS_CITZ_sign_date.xlsx'

#ISOCodes<-'https://datahub.io/core/country-list/data.csv'
GET(URL,write_disk(tf<-tempfile(fileext = '.xlsx')))

SP<-read_excel(tf,1L,skip = 4)
```

Identifying data source date ranges based on title within file.
```{r}
SPDates<-read_excel(tf,1L,col_names = FALSE)[1,1]
print(SPDates)
SPDates<-SPDates[[1]]

#Reg ex to pull date ranges

StartDate<-'+\\d{4}'

#Date Ranges 
Years<-regmatches(SPDates,gregexpr(StartDate,SPDates,
                                     ignore.case = TRUE))[[1]]
#Range of years within the dataset
YearsRange<-Years[1]:Years[2]

```

### Cleaning Column Names

```{r}

#Columns to clean 
SPColnames<-colnames(SP)[2:ncol(SP)]

#Add in years to column names
replicated_years<-rep(YearsRange,each = round(length(SPColnames)/length(YearsRange),0)+1)
#At this point each year in the replicated years variable represents either a month, a quarter or a year total. The next line of code will keep only dates from the start of the data source to the most recent month. 
replicated_years<-replicated_years[1:length(SPColnames)]
print(replicated_years)
SPColnames<- paste(SPColnames,replicated_years, sep = '-')
SPColnames
#Drop extra characters
SPColnames<-sub("\\..*?-","-",SPColnames)
#Add total title to appropriate column names. 
SPColnames<-ifelse(nchar(SPColnames)==5, paste("Total", SPColnames,sep = ''), SPColnames)
#Apply titles to DS
colnames(SP)<-c('Country',SPColnames)
```

cleaning some country codes 
	
Bahama Islands, The = Bahamas
Saint-Martin = Saint Martin (French Part)
Stateless, Country of citizenship not stated = 'Other' END

```{r}
SP$Country<-ifelse(SP$Country=='Bahama Islands, The','Bahamas',
                     ifelse(SP$Country=='Saint-Martin','Saint Martin (French part)',
                            ifelse(SP$Country=='Netherlands Antilles, The' | SP$Country=='Kosovo, Republic of'| SP$Country=='Stateless'|SP$Country=='Country of Citizenship not stated','Other',SP$Country)))
```

## Cleaning Missing Values
* For now we will transform the '--' character to the average/mean of 1-4

```{r}
##We want to remove invalid rows, drop everything after the first NA
firstNA<-which(is.na(SP[,1]))[1]
SP<-subset(SP,seq_len(nrow(SP))<firstNA)

#anything with -- will be the average of 1:4
SP[SP=='--']<-as.character(mean(c(1,4)))
```

###Pivot
* Lets turn the data set from a wide to long format

```{r}
#Convert columns to integer

SP<-mutate_at(SP,vars(colnames(SP[2:length(SP)])),
               function(x) as.integer(gsub(',',"",x)))

SP<- SP %>%
  pivot_longer(
    cols = colnames(SP[2:ncol(SP)]),
    names_to = c('Period'),
    values_to = 'Total Study Permits'
  )

```

* Creating a field to identify the level/time period detail for each record. 
```{r}
SP$Level<-ifelse(substr(SP$Period,1,1)=='Q','Quarterly',
                 ifelse(substr(SP$Period,1,1)=='T','Annually','Monthly'))
```

*Adding in country codes. Going to be used to bring in regions.

```{r}
#install.packages('countrycode')
library('countrycode')

#Iso 3 code
SP$CountryISO<-countrycode(sourcevar=SP$Country,origin = 'country.name',
            destination = 'iso3c'
            )

#continent 
SP$Continent<-countrycode(sourcevar=SP$Country,origin = 'country.name',
            destination = 'continent'
            )
#World Bank Region 
SP$WorldBankRegion<-countrycode(sourcevar=SP$Country,origin = 'country.name',
            destination = 'region'
            )


#Could look at correlation between study permits and different datasets on world bank

```

* Next steps could be bring in data for the other 3 types of temporary residents (work permits, visitor rcords and temp residents.

https://www.canada.ca/en/immigration-refugees-citizenship/services/new-immigrants/status-documents.html

work permits: https://open.canada.ca/data/en/dataset/360024f2-17e9-4558-bfc1-3616485d65b9

Temporary Work Permits: https://open.canada.ca/data/dataset/360024f2-17e9-4558-bfc1-3616485d65b9/resource/2e910f7d-100b-49d4-bab4-a32922af8459


```{r}
#For the purpose of the dashboard I will only focus on months and remove unique person count. 
SP<- SP %>% 
  filter(Level=='Monthly' & Country!='Total unique persons')

## Cleaning up contient and region field
NA_Country_Cont<-unique(SP[is.na(SP$Continent),][,1])
SP$Continent<-ifelse(SP$Country=='United States Minor Outlying Islands','Americas',
                      ifelse(is.na(SP$Continent)==TRUE,'Other',SP$Continent))

unique(SP$Continent)
NA_Country_Region<-unique(SP[is.na(SP$WorldBankRegion),][,1])
SP$WorldBankRegion<-ifelse(SP$Country=='Reunion','Sub-Saharan Africa',
                            ifelse(SP$Country=='United States Minor Outlying Islands','Other',
                                   ifelse(SP$Country=='Western Sahara','Sub-Saharan Africa',
                                          ifelse(is.na(SP$WorldBankRegion) ==TRUE,'Other',SP$WorldBankRegion)
                                          )))
unique(SP$WorldBankRegion)
```


```{r}
write.csv(SP,'C:\\Users\\User\\Desktop\\Tableau\\International\\TempResPerHold\\StudyPermits.csv')


```

